---
title: "New Test"
author: "Aditya Wakade"
date: "2 March 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the data into csv from Google drive download
Find summary of all columns


```{r}
setwd("C:/Users/wakad/Documents/INFX 573/project")
fomc1 = read.csv("Prod2.csv")
fomc1 = fomc1[,c(-1,-8,-11,-15)]

summary(fomc1)
nrow(fomc1)
```
We aim to find out factors that affect the final Bill Amount of an order. Based on our findings we aim to optimize values for maximum bill amount.




We find the central tendency of the Bill Amount

```{r}
hist(fomc1$Bill.Amount)
```
Mean lies at 966



Looking at the data, mean ~ 1000 and median is ~ 800, we see that this is a J shaped histogram so the median is biased to the left side of the histogram as the histogram is skewed to right.

Add a column with Bill amount notation greater than mean 966
Define a new column which holds notation for Metropolitan cities of India

```{r}
fomc1$Billnot <-rep(0, nrow(fomc1))
fomc1$Billnot[fomc1$Bill.Amount > 966] <-1
fomc1$Citymet <-rep(0, nrow(fomc1))
fomc1$Citymet[fomc1$City == "Mumbai" | fomc1$City == "Bengaluru" | fomc1$City == "Delhi" | fomc1$City == "Kolkata" | fomc1$City == "pune" | fomc1$City == "Chennai" | fomc1$City == "Hyderabad"] <-1
focmtrain = fomc1[1:5000,]
fomctest = fomc1[5001:6000,]
```

We use the factors of Category of items, price range of items, State of Delivery, Quantity,
 City metropolitan For finding how they affect the Bill Amount being above or below 966
Finding binomial general linear model...as.character(pricecateg)+as.factor(Quantity)+

```{r}
mod1 <-glm(Billnot~as.factor(Month)+geogloc+ as.character(Citymet)+as.character(Quantity)+ as.character(pricecateg),data=fomc1[,!colnames(fomc1)%in%c("Bill.Amount")], family="binomial")
summary(mod1)
```
Observations

with *** statistical significance

The log odds of the bill amount being above 966 reduces because of -2.6386 for Dry Fruits
and -0.99 because of Sweets

others

the log odds of bill being above 966 reduces by -3.49 because of juice and -0.46 because of snacks

Being a metropolitan city weakly affects the bill amount being more than 966 

sPICES WEAKLY indicate bill more than 966
 
 
Other observations like Ready to eat being for higher bill have very high standard error


We try and find the effect of price points of products on Total bill amount


```{r}

x<-fomc1$Price
plot(fomc1$Billnot~x, col="blue", xlab = "Price of Product", ylab = "Bill Notation")
fits <-fitted(mod1)
points(x, fits, pch = 4 , cex=0.3)
abline(0.5, 0)
```
Hence products between 100-200 rs boost the chances of total bill being above 966


Next we try to find Quantity measure which is optimal for higher bill amounts
```{r}
x <- fomc1$Quantity
plot(fomc1$Billnot~x, col="blue", xlab = "Quantity", ylab = "Bill Notation")
fits <-fitted(mod1)
points(x, fits, pch = 4 , cex=0.3)
abline(0.5, 0)
```


Hence, quantities 8 and 9 closely followed by 7 provide larger probability of the bill being higher than 966


We find the error percentage of our logistic model

lowest percentage error
```{r}
tab <-table(fomc1$Billnot, fits>=0.5)
(tab[1,2]+tab[2,1])/sum(tab)
```
Thats about 5% error


We find the AUC metric
```{r}
predictpr = predict(mod1, type = c("response"))
library(pROC)
library(AUC)
y <-factor(fomc1$Billnot)
rr <-roc(y~predictpr)
plot(rr)
auc(rr)
```

```{r}
library(boot)
cv.glm(glmfit = mod1, data = na.omit(fomc1[,c('Month', )]), K = 20)
```
```{r}
pr = predict(mod1, newdata = fomctest, type = c("response"))
library(AUC)
library(pROC)
y <-factor(fomctest$Billnot)
rr <-roc(y,pr)
plot(rr)
auc(rr)
```



We divide the whole dataset into metropolitan and non metropolitan cities

```{r}
metrodata = fomc1[(fomc1$Citymet == 1),]
nonmetro = fomc1[(fomc1$Citymet == 0),]
nrow(metrodata)
nrow(nonmetro)
```


We check the frequency of categories in metro and non metro. However, since total orders in the two subsets are uneven as seen in the previous tab, we find a normalization ratio which is 4880/1120 = 4.357143
```{r}
test = xtabs(~Category+Citymet, fomc1)
test[,2] = test[,2]*4.357143
test
```
Organic foods are peferred by non metro cities  (find other observations)




MOst fequent items in metro
```{r}
metrodata$Category = factor(metrodata$Category, levels(metrodata$Category))
barplot(table(metrodata$Category))
#text(metrodata$Category, label = levels(droplevels(metrodata$Category)))
```

Finding frequent 20 items in metro

```{r}
x = sort(-table(metrodata$Item.Name))
head(x,20)

```

most frequent items in non metro
```{r}
barplot(table(nonmetro$Item))
text(nonmetro$Item, label = levels(droplevels(nonmetro$Item.Name)))
```

frequent 20 in non metro
```{r}
x = sort(-table(nonmetro$Item.Name))
head(x,20)
```
Find common items between metro and non metro to define overall popular items
